---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# slidetool

<!-- badges: start -->
<!-- badges: end -->

Retrieve Data From Whole Slide Images in R

## Overview

`slidetool` is an R package that leverages OpenSlide to interact with whole slide images, also known as virtual slide files.
It provides functions to extract metadata, list associated images, read specific regions of slides, and more.

## Installation

### Installing the development version

You can install the development version of slidetool from [GitHub](https://github.com/) with:

```{r eval=FALSE}
remotes::install_github("luciorq/slidetool")
# or
pak::pkg_install("github::luciorq/slidetool")
```

## Usage

```{r}
library(slidetool)
```

Next, install OpenSlide dependencies using the `install_slidetool()` function:

```{r}
install_slidetool()

slidetool("--version") |>
  _$stderr |>
  message()
```

List associated images for a slide:

```{r}
# Replace with the path to your slide file
slide_path <- fs::path_package("slidetool", "extdata", "CMU-1-Small-Region.svs")

associated_images <- slidetool_assoc_list(slide_path)
print(associated_images)
```

Read Thumbnail from a SVS file:

```{r}
thumb_path <- slidetool_assoc_read(slide_path, assoc = "thumbnail")

img_obj <- EBImage::readImage(thumb_path)
EBImage::display(img_obj, method = "raster")
```

Extract a specific region from a slide and save it as a PNG image:

```{r}
# Define parameters
x_coord <- 1000
y_coord <- 1000
level <- 0
width <- 500
height <- 500
# output_file <- fs::file_temp(pattern = "region", ext = "png")

# Read the region
region_path <- slidetool_region_read(
  path = slide_path,
  x = x_coord,
  y = y_coord,
  level = level,
  width = width,
  height = height
)

# Check if the file was created
if (file.exists(region_path)) {
  message("Region saved to ", region_path)
}
```

```{r}
img_obj <- EBImage::readImage(region_path)
print(img_obj)
```

```{r}
EBImage::display(img_obj, method = "raster")
```

Get all OpenSlide properties and their values:

```{r}
properties <- slidetool_prop_list(slide_path)
print(names(properties))
print(properties["openslide.level-count"])
```

Retrieve the value of a specific property:

```{r}
property_name <- "openslide.vendor"
property_value <- slidetool_prop_get(slide_path, property_name)
print(paste0(property_name, ": ", property_value))
```

```{r}
# magick::image_read(region_path)
```


## License

This project is licensed under the GNU AFFERO GENERAL PUBLIC LICENSE v3+.
See the [LICENSE](./LICENSE.md) file for details.

## Contributing

Contributions are welcome! Please submit issues and pull requests on the [GitHub repository](https://github.com/luciorq/slidetool).

---

## Aknowledgments

- [OpenSlide](https://openslide.org/) for providing a reliable interface to read whole slide images.
- [condathis](https://github.com/luciorq/condathis) for environment management.
